An Extensible Approach to Verification of Embedded Systems

### 2014-05-01 tech-project/abstract.tex
changes since initial turn in

In this paper we present a flexible means of specifying and verifying the correctness of software for embedded network systems. Our approach uses RESOLVE, an imperative, component based programming and mathematical specification language, to verify the functional correctness of embedded applications. In doing so, we enrich the work originally presented in \cite{regula:2010} with the following: A model view controller (MVC) based implementation of a RESOLVE to C translator, a dynamic memory allocation scheme tailored towards embedded systems running the code generated by our tool, and the addition of a new language keyword that enables users to pair custom RESOLVE specifications with `externally' implemented (non-native RESOLVE) realizations. We demonstrate these additions on an LED (Light emitting diode) driver that showcases recent mathematical developments, as well as formal verification of a toggling capability enhancement that we demonstrate running on a Telos mote.
\end{abstract}

>>>>

In this paper we present a flexible means of specifying and verifying the correctness of software for embedded network systems. Our approach uses RESOLVE, an imperative, component based programming and mathematical specification language, to verify the functional correctness of embedded applications. In doing so, we enrich the work originally presented in \cite{regula:2010} with the following: A model view controller (MVC) based implementation of a RESOLVE to C translator, a dynamic memory allocation scheme tailored towards embedded systems running the code generated by our tool, and the addition of a new language keyword that enables users to pair custom RESOLVE specifications with `externally' implemented (non-native RESOLVE) realizations. We demonstrate these additions on an LED (Light emitting diode) driver that showcases recent mathematical developments, as well as formal verification of a toggling capability enhancement.
\end{abstract}

### 2014-05-01 tech-project/implementation.tex
updated the demo tex
changes since initial turn in

% IN THIS SECTION SHOW SOME OF OUR TRANSLATED CODE... AND WALK THROUGH IT THE SAME WAY YOU DID WITH THE CONCEPT IN RESOLVE. BUT BEFORE YOU DO THAT, SHOW THE PICTURE (The one thats already here showing high level relationships and update it!).
\section{Implementation}
Development of our C translation tool can be logically partitioned into three distinct phases: 
\begin{enumerate}
\item Arriving at a translation model (or, strategy) for an accurate C representation of RESOLVE.
\item Implementing reusable mechanisms for carrying out the C code generation process.
\item Creation of a memory manager capable of safely allocating and freeing dynamic memory required by the generated code.
\end{enumerate}
We conclude this section with a demonstration of each of these phases working in tandem on the LED component discussed in Section \ref{sec:specifiying}.

\subsection{C Translation Model}
One of the primary challenges in translating from RESOLVE to C is finding a suitable C analog for each RESOLVE module and the constructs allowable in each. Indeed, since we are dealing with an environment where functional correctness is a primary concern, it is important that the code generated by our tool represents as closely as possible the original RESOLVE source. In an effort to make such considerations, at the highest level, the C code we generate makes special considerations for concepts, facilities, and realizations. This scheme is depicted in Figure \ref{fig:relationship} and briefly summarized in Sections \ref{sec:conceptoverview} - \ref{sec:facilitiesrealizations}.

\begin{figure}
\begin{center}
\includegraphics[scale=.60]{figs/relationship.pdf}
\end{center}
\caption{Relationship between RESOLVE module types and the C code generated from each.}
\label{fig:relationship}
\end{figure}

@@

%The memory model must also be considered as an additional verifiable component. Currently, RESOLVE is not capable of creating a complete specification of memory\footnote{RESOLVE is an object based language and has variable sized structures. Using dynamically allocatable memory is the favored approach to allow arbitrarily sized data.}. Thus, a memory model must be realized without a specification. To provide a straight forward translation from RESOLVE to C, we provide a dynamically-based memory allocator for use on embedded systems.

%Unsurprisingly, concepts are represented in C by a single header .h listing the various methods

\begin{figure*}
\begin{center}
\includegraphics[scale=.55]{figs/ast_traversal2.pdf}
\end{center}
\caption{The general flow of information from the AST (first), to user defined templates (middle), ending with formed output (last).}
\label{fig:ast}
\end{figure*}

@@

Shown in Figure \ref{fig:ast} is a high level depiction of steps taken in translating this operation to C. The first box depicts the AST of \texttt{Nothing}, where nodes are represented with boxes labeled by the constructs they contain. Throughout the walk of this tree, useful information (such as the operation's name, ``Nothing") are extracted from the nodes containing these constructs, and added to a user defined string-template\footnote{A template can simply be thought of as a ``document with holes" which the user choses when and how to fill.}, in this case: \texttt{c\_function\_def}.

@@

That is, the only actual work being performed within the C translator is forwarding information gathered from individual treenodes, to a series of externally defined templates. This allows us to exploit (in design pattern parlance) a strict model view controller (MVC) separation in the translator's codebase between the mechanism that does the AST visiting (controller), the tree nodes from which we're adding information to templates (model), and the external file containing all available C language templates which shape our output (view).

We feel this approach lends itself well to the challenge discussed in this paper, as this separation allows us to easily iterate changes to our generated C code without needing to concern ourselves with the Java written inside the compiler itself. This allows us to easily and iteratively tweak generated code -- making arbitrarily complicated changes and optimizations, some of which are detailed later in Section \#.

%%%%%%%%%%%%%%%%%%%%%%%
%           Memory Allocation Section              %
%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Memory Allocation}\label{sec:mem}

Any model seeking to convert RESOLVE to a lower level representation such as C demands a strategy for handling any memory utilized by the generated code. While dynamic allocation is typically the norm, it is not however the first choice for embedded applications. With the Telos mote constrained to 128 bytes of RAM, developers targeting embedded platforms tend to favor static memory allocation over dynamic, due the memory efficiency it affords.

@@

In spite of this, our reasoning for choosing to pursue dynamic allocation over static, rests in the language itself. Being a language that relies heavily on the notion of formally specified objects (components), everything in RESOLVE from the `primitive' types such as booleans and integers, to the more complicated ones such as Stacks and Queues, are modeled, specified, and implemented in the same way: Using the standard RESOLVE machinery discussed in Section \ref{sec:specifiying}.

As such, even the simplest RESOLVE programs still rely heavily on the notion of objects coming and going. Thus, from a verification perspective, creating a tool that allows our translated code to create and destroy such objects in the manner the language itself uses is an important component to capture in formally verified code.

\subsubsection{Allocation using \texttt{salloc}}

The particular allocation scheme that we chose to use is \texttt{salloc()}: A first fit memory allocator. Rather than conventional, heap-based allocators, \texttt{salloc()} uses stack memory for allocation. This approach requires that a fixed size of memory is chosen at compilation. In addition, \texttt{salloc()}, requires a section of meta-data, which is denoted as a \texttt{block}, for each record in the memory pool. A \texttt{block} holds referential information of neighboring blocks, the size of the record the block maintains, and if the block is free or is in use. Figure~\ref{fig:stack} is an example of the stack after several calls to \texttt{salloc()}.

\begin{figure}[!htb]
%\centering
\includegraphics[scale=.55]{figs/stack.pdf}
\caption{Representation of memory using salloc()}
Memory allocated with salloc() allocates a block in front of data allocated. Blocks point to their immediate neighbors and hold their size and usage.
\label{fig:stack}
\end{figure}

@@

A memory allocator must provide a mechanism to release, or free memory in order to indicate that it is not being used and can be reallocated. The \texttt{sfree()} function provides deallocation for memory allocated with \texttt{salloc()}. Figure~\ref{fig:free} shows an example set of memory before and after calling \texttt{sfree()}.

\begin{figure}[!htb]
%\centering
\includegraphics[scale=.55]{figs/sfree.pdf}
\caption{sfree to free memory}
\label{fig:free}
\end{figure}

@@

A common problem that can occur in memory allocation is fragmentation. During execution, a program can allocate and deallocate an arbitrary number of times. Figure~\ref{fig:fragmentation} shows problems that result from this. This problem is magnified on embedded systems due to their limited memory capacities. There are simple optimizations that can reduce fragmentation.

Block splitting, as shown in Figure~\ref{fig:split}, occurs during allocation. It allows blocks of greater size to be partitioned into the size requested by the allocator.

Fusing blocks is another technique used when \texttt{sfree()} is called. When memory is deallocated, neighboring free blocks are coalesced to form a single large block, as shown in Figure~\ref{fig:fuse}.

\begin{figure}[!htb]
\centering
\includegraphics[scale=.55]{figs/fragmentation.pdf}
\caption{Performing naive allocation and deallocations, allocatable blocks are restricted to those that match the exact size of memory requested from the allocator}
\label{fig:fragmentation}
\end{figure}

>>>>

% IN THIS SECTION SHOW SOME OF OUR TRANSLATED CODE... AND WALK THROUGH IT THE SAME WAY YOU DID WITH THE CONCEPT IN RESOLVE. BUT BEFORE YOU DO THAT, SHOW THE PICTURE (The one thats already here showing high level relationships and update it!).
\section{Implementation}\label{sec:impl}
Development of our C translation tool can be logically partitioned into three distinct phases: 
\begin{enumerate}
\item Arriving at a translation model (or, strategy) for an accurate C representation of RESOLVE.
\item Implementing suitable mechanisms for carrying out the C code generation process.
\item Creation of a memory manager capable of safely allocating and freeing dynamic memory required by the generated code.
\end{enumerate}
We conclude this section with a demonstration of each of these phases working in tandem on the LED component discussed in Section \ref{sec:specifiying}.

\subsection{C Translation Model}
One of the primary challenges in translating from RESOLVE to C is finding a suitable C analog for each RESOLVE module and the constructs allowable in each. Indeed, since we are dealing with an environment where functional correctness is a primary concern, it is important that the code generated by our tool represents as closely as possible the original RESOLVE source. In an effort to make such considerations, at the highest level, the C code we generate makes special considerations for concepts, facilities, and realizations. This scheme is depicted in Figure \ref{fig:relationship} and briefly summarized throughout Sections \ref{sec:conceptoverview} - \ref{sec:facilitiesrealizations}.

\begin{figure}
\begin{center}
\includegraphics[scale=.55]{figs/relationship.pdf}
\end{center}
\caption{Relationship between RESOLVE module types and the C code generated from each.}
\label{fig:relationship}
\end{figure}

\begin{figure*}[!htb]
\begin{center}
\includegraphics[scale=.54]{figs/ast_traversal2.pdf}
\end{center}
\caption{The general flow of information from the AST (first), to user defined templates (middle), ending with formed output (last).}
\label{fig:ast}
\end{figure*}

@@

@@

Shown in Figure \ref{fig:ast} is a high level depiction of steps taken in translating this operation to C. The first box depicts the AST of \texttt{Nothing}, where nodes are represented as boxes labeled by the constructs they contain. Throughout the walk of this tree, useful information such as the operation's name, ``Nothing", are extracted from nodes, and added as parameters to user defined templates\footnote{A template can simply be thought of as a ``document with holes" which the user choses when and how to fill.}, in this case: \texttt{c\_function\_def}.

@@

That is, the only actual work being performed within the C translator is forwarding information gathered from individual treenodes, to a series of externally defined templates. This allows us to exploit (in design pattern parlance \cite{krasner:1988}) a strict model view controller (MVC) separation in the translator's codebase between the mechanism that does the AST visiting (controller), the tree nodes from which we're adding information to templates (model), and the external file containing all available C language templates which shape our output (view).

We feel this approach lends itself well to the challenge discussed in this paper, as this separation allows us to easily iterate changes to our generated C code without needing to concern ourselves with the Java written inside the compiler itself.

%%%%%%%%%%%%%%%%%%%%%%%
%           Memory Allocation Section              %
%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Strategy for Dynamic Memory Allocation}\label{sec:mem}

Any model seeking to convert RESOLVE to a lower level representation such as C demands a strategy for handling memory utilized by the generated code. While dynamic allocation is typically the norm, it is not however the first choice for embedded applications. With the Telosb mote constrained to 128 bytes of RAM, developers targeting embedded platforms tend to favor static memory allocation over dynamic, due the memory efficiency it affords.

@@

In spite of this, our reasoning for choosing to pursue dynamic allocation over static, rests in the language itself. Being a language that relies heavily on the notion of formally specified, verifiable objects (components), everything in RESOLVE from the `primitive' types such as booleans and integers, to the more complicated ones such as Stacks and Queues, are modeled, specified, and implemented in the same way: Using the standard component machinery showcased in Section \ref{sec:specifiying}.

As such, even the simplest RESOLVE programs still rely heavily on the notion of objects coming and going. Thus, from a verification perspective, creating a tool that allows our translated code to create and destroy such objects in a manner consistent with the language is an extremely important aspect towards maintaining the ``correct-by-construction" property of our automatically generated C code.

The previous iteration of this tool \cite{regula:2010} (favoring static allocation) carries out ``pseudo allocation"  by effectively introducing second pass to the compiler dedicated to replacing all formal parameters, with their corresponding facility-instantiated literal values. This additional, hidden step to compilation is less than ideal in the context of a verifying compiler designed from the ground up to be a single pass system. We feel the work we present here avoids these wrinkles with a dynamic approach.

\subsubsection{Allocation using ``salloc''}

The dynamic allocatior we create for our generated code is \texttt{salloc()}: A first fit memory allocator. Rather than conventional, heap-based allocators, \texttt{salloc()} uses stack memory for allocation. This approach requires that a single fixed block of memory is chosen at compilation. In addition, \texttt{salloc()}, requires a section of meta-data, denoted as a \texttt{block}, for each record in memory. A \texttt{block} holds referential information about neighboring blocks, the size of the record the block maintains, and if the block is free or is in use. Figure~\ref{fig:stack} is an example of the stack after several calls to \texttt{salloc()}.

\begin{figure}[!htb]
%\centering
\includegraphics[scale=.55]{figs/stack.pdf}
\caption{Memory allocated with salloc() allocates a block in front of data allocated. Blocks point to their immediate neighbors and hold their size and usage.}

\label{fig:stack}
\end{figure}

@@

A memory allocator must provide a mechanism to release, or free a section of memory in order to indicate that it is not being used and that it may be reallocated. Our \texttt{sfree()} function, which provides this functionality, is illustrated in Figure~\ref{fig:free} -- depicting memory before and after a call to \texttt{sfree()}.

\begin{figure}[!htb]
%\centering
\includegraphics[scale=.55]{figs/sfree.pdf}
\caption{An illustration of sfree.}
\label{fig:free}
\end{figure}

@@

A common problem that can occur in memory allocation is fragmentation. During execution, a program can allocate and deallocate memory an arbitrary number of times. Figure~\ref{fig:fragmentation} depicts some of the problems that might arise from this such as a decreased success rate for allocation and increased overall memory usage. The problem is especially magnified on embedded systems, due to their limited memory capacity. We employ two simple optimizations to help reduce this.

\begin{itemize}
\item \textbf{Block splitting:} As shown in Figure~\ref{fig:split}, unused blocks can be repartitioned to the size of memory needed by \texttt{salloc()}.
\item \textbf{Block fusing:}
This technique is applied during deallocation. When memory is deallocated by \texttt{sfree()}, neighboring free blocks containing unused memory are coalesced into a single large block, as shown in Figure~\ref{fig:fuse}.
\end{itemize}

%The first is block splitting, as shown in Figure~\ref{fig:split}, occurs during allocation. It allows blocks of greater size to be partitioned into the size requested by the allocator.

%Fusing blocks is another technique used when \texttt{sfree()} is called. When memory is deallocated, neighboring free blocks are coalesced to form a single large block, as shown in Figure~\ref{fig:fuse}.

\begin{figure}[!htb]
\centering
\includegraphics[scale=.55]{figs/fragmentation.pdf}
\caption{Performing naive allocation and deallocations restricts allocatable blocks to those that match the size requested.}
\label{fig:fragmentation}
\end{figure}

### 2014-05-01 tech-project/introduction.tex
changes since initial turn in

\section{Introduction}
\label{sec:intro}
Within little more than a decade, the area of embedded network systems and wireless sensing has exploded in popularity within industry and academia alike. Tempering however this extremely quick rise in popularity is the inherent difficulties in developing applications that function as intended in low power, event-driven environments. In response to these difficulties, a variety of tools and languages have been put forth  to help ease the burden on developers. On one end of this effort are languages such as NesC, (Network embedded sensor C) which strive to minimize concurrency issues and other common sources of error by hiding libraries of pre-written drivers underneath hierarchies of software and interface level abstractions. The other end of this effort is largely comprised of simulation tools such as TOSSIM, Cooja, and Arora that make use of high-fidelity simulations to model networks offline in controlled, repeatable environments. Though these and other tools have indeed proven invaluable in allowing users to test and reason about event-driven code prior to deployment, they remain incapable of providing complete assurance that code will behave as expected when deployed in the field.

We approach this problem by using RESOLVE (Reusable SOftware Language with VErification) as means of authoring, specifying, and ultimately verifying code for embedded network systems. Our decision to use RESOLVE as a language frontend -- as opposed to verifying C code directly -- ultimately stems from a verification amenability standpoint: Not only does RESOLVE prohibit verification crippling operations such as uncontrolled referencing and aliasing (prevalent in C and many other current languages)\cite{kulczycki:2004}, but also embodies a number of other characteristics ideal for embedded platforms including:

\begin{itemize}
\item \textbf{Modularity} RESOLVE enforces a strict separation of concerns between module specifications and client implementations. As a result, for any one particular specification, there can be any number of interchangeable implementations. This separation is ideal considering that many embedded applications happen to fit this pattern nicely: Various drivers oftentimes provide a common set of functionality, but in general have many distinct implementations that vary arbitrarily from platform to platform, vendor to vendor.

\item \textbf{Mathematical flexibility} RESOLVE offers a rich, mathematical type system that allows users to either draw from a library of preexisting mathematical units when writing specifications, or simply create their own. This is ideal in a setting where drivers might encompass a wide spectrum possible applications, where each might require unique, new mathematical developments.% reword this to be less vague.

@@

The paper is organized as follows: First, we open with a brief overview of the Telos mote platform. Next, we present revised specifications of an LED driver component with a formally verified enhancement. This section is concluded with a review of verification results, and a discussion of any relevant theorems and verification conditions (VCs) used. The remainder of the paper is spent detailing the model of C code generated by the tool, giving a quick overview of generation process itself, and detailing a dynamic, stack-based memory allocation tool utilized by the translated code. We conclude with suggestions for tool improvements, as well as a review of some longer term research goals.

>>>>

\section{Introduction}
\label{sec:intro}
Within little more than a decade, the area of embedded network systems and wireless sensing has exploded in popularity within industry and academia alike. With this quick and widespread proliferation of embedded devices, the need for provably correct embedded code increases -- especially in those industries where the cost of failure could be extreme both in terms of cost and human life.

From a non-verification related perspective, a variety of tools and languages have been put forth to help make developing embedded applications a simpler, less error prone endeavor. On one end of this effort are languages such as NesC, (Network embedded sensor C) which strive to minimize concurrency issues and other common sources of error by hiding libraries of pre-written drivers underneath hierarchies of software and interface level abstractions. The other end of this effort is largely comprised of simulation tools such as TOSSIM, Cooja, and Avrora that make use of high-fidelity simulations to model networks offline in controlled, repeatable environments. Though these and other tools have indeed proven invaluable in allowing users to test and reason about event-driven code prior to deployment, they remain incapable of providing complete assurance that code will behave as expected when deployed.

We approach this problem by using RESOLVE (Reusable SOftware Language with VErification) as means of authoring, specifying, and ultimately verifying code, prior to execution, for embedded network systems. Our decision to use RESOLVE as a language frontend -- as opposed to verifying C code directly -- ultimately stems from a verification amenability standpoint: Not only does RESOLVE prohibit verification crippling operations such as uncontrolled referencing and aliasing (prevalent in C and many other current languages)\cite{kulczycki:2004}, but also embodies a number of other characteristics ideal for embedded platforms including:

\begin{itemize}
\item \textbf{Modularity} RESOLVE enforces a strict separation of concerns between module level specifications and client level implementations. As a result, for any one particular specification, there can be any number of interchangeable implementations. This separation is ideal considering that many embedded applications happen to fit this pattern nicely: Various drivers oftentimes provide a common set of functionality, but in general have many distinct implementations that vary arbitrarily from platform to platform, vendor to vendor.

\item \textbf{Mathematical flexibility} RESOLVE offers a rich, mathematical type system that allows users to either draw from a library of pre-existing mathematical units when writing specifications, or simply create their own. This is ideal in a setting where drivers might encompass a wide spectrum possible applications -- each requiring new mathematical developments.

@@

The paper is organized as follows: First, we open with a brief overview of the Telos mote platform. Next, we present revised specifications of an LED driver component with a formally verified enhancement. This section is concluded with a review of verification results, and a discussion of any relevant theorems and verification conditions (VCs) used. The remainder of the paper is spent detailing the model of C code generated by the tool, giving a quick overview of generation process itself, and detailing a dynamic, stack-based memory allocation tool utilized by the translated code. We conclude with suggestions for tool improvements, as well as a review of some long term research goals.

### 2014-05-01 tech-project/led_trans.tex
changes since initial turn in

% An LED Demo Translation
\section{LED Demo}

@@

#include ".../RESOLVE.h"
#include ".../Facilities/.../Std_Bool_Fac.h"
#include ".../Facilities/.../Std_Int_Fac.h"
#include ".../Facilities/.../Std_Clock_Fac.h"
#include ".../Concepts/.../LED_Template.h"
#include ".../Concepts/.../Std_LED_Realiz.h"
#include ".../Concepts/.../Toggling_Realiz.h"

@@

void LED_Telos_Demo_create();
void LED_Telos_Demo_destroy();
#endif
\end{verbatim}
\end{minipage}
\hspace{1.5cm} 
\begin{minipage}{0.4\textwidth}
\begin{verbatim}
/*
 * Generated by the RESOLVE to C translator. 
 * This file should not be modified.
 */
#include "LED_Telos_Demo.h"

void LED_Telos_Demo_create() {
  Std_Bool_Fac_create();
  Std_Int_Fac_create();
  Std_Clock_Fac_create();
  r_type_ptr __arg_0 = Std_Int_Fac_Var.core->
    createFromInt(4, Std_Int_Fac_Var.core->Integer);
  LED_Fac_Var.core = 
    new_Std_LED_Realiz_for_LED_Template(__arg_0);
  LED_Fac_Var.Toggling_Capability = 
    new_Toggling_Realiz_for_
    Toggling_Capability_of_LED_
    Template(LED_Fac_Var.core);
  Std_Int_Fac_Var.core->Integer->
    destroy(__arg_0, Std_Int_Fac_Var.core->Integer);
}

void LED_Telos_Demo_destroy() {
    free_Std_LED_Realiz_for_LED_Template(
    				    LED_Fac_Var.core);
    Std_Bool_Fac_destroy();
    Std_Int_Fac_destroy();
}
        ...
\end{verbatim}
\end{minipage}
\end{figure*}

>>>>

\section{Putting It All Together}

@@

#include ".../RESOLVE.h"
  ...
#include ".../Concepts/.../LED_Template.h"
#include ".../Concepts/.../Std_LED_Realiz.h"
#include ".../Concepts/.../Toggling_Realiz.h"

@@

void LED_Telos_Demo_create();
void LED_Telos_Demo_destroy();
#endif
\end{verbatim}
\end{minipage}
\hspace{1.5cm} 
\begin{minipage}{0.4\textwidth}
\begin{verbatim}
#include "LED_Telos_Demo.h"

void LED_Telos_Demo_create() {
  Std_Bool_Fac_create();
  Std_Int_Fac_create();
  Std_Clock_Fac_create();
  r_type_ptr __arg_0 = Std_Int_Fac_Var.core->
    createFromInt(3, Std_Int_Fac_Var.core->Integer);
  LED_Fac_Var.core = 
    new_Std_LED_Realiz_for_LED_Template(__arg_0);
  LED_Fac_Var.Toggling_Capability = 
    new_Toggling_Realiz_for_Toggling_Capability_
    of_LED_Template(LED_Fac_Var.core);
  Std_Int_Fac_Var.core->Integer->
    destroy(__arg_0, Std_Int_Fac_Var.core->Integer);
}

void LED_Telos_Demo_destroy() {
    free_Std_LED_Realiz_for_LED_Template(
    				    LED_Fac_Var.core);
    Std_Bool_Fac_destroy();
    Std_Int_Fac_destroy();
} ...
\end{verbatim}
\end{minipage}
\caption{C Translation of \texttt{LED\_Telos\_Demo}.}
\label{fig:demo}
\end{figure*}

Given our RESOLVE realization of the \texttt{LED\_Telos\_Demo} described in Section 3, we now discuss our translator's output in greater detail.

Depicted in figure \ref{fig:demo} is the C translation of \texttt{LED\_Telos\_Demo} consisting of a .h (left) and a portion of the .c (right). The \texttt{LED\_Telos\_Demo\_create()} function provides instantiations for all global facilities used in \texttt{LED\_Telos\_Demo}, including those for standard types such as integers, booleans, clock, as well as local facilities such as \texttt{Leds\_Fac}. One additional thing to note within this create method is the temporary \texttt{\_\_arg\_0} integer variable -- representing the formal value \texttt{Strip\_Length} found within the concept. Note that this variable does not appear in the original source, but rather, is created by the translator in an effort to keep track of all literal argument values that require manual deletion.

The \texttt{Leds\_Fac} facility from the original source, represented here as a struct, uses a \texttt{core} field to provide a means of accessing the methods and types in the original concept, while any remaining fields correspond to enhancement pointers that serve a similar end -- in this case, providing access to the operations declared within  \texttt{Toggling\_Capability}.

Clearly facilities and other modules highly reliant on facility instantiations are no doubt going to consume much more memory than the small example detailed here. To help alleviate some of th

%\texttt{new\_Toggling\_Realiz\_for\_Toggling\_Capability\_of\_LED\_Template()}. Similarly, this initialization and allocation step must occur for the Integer, Boolean, and Clock facilities. From this relatively small example of LED toggling from RESOLVE to C,  complex programs will require even more attention to amount of memory used for each concept and facility realization. Curtailing the management of this memory by placing the responsibility on an allocator provides a key motivation for using \texttt{salloc()} and \texttt{sfree()}.

%As seen from above, RESOLVE to C translations can produce verbose output. The code presented to the right introduces pseudocode for a toggling capability of an LED strip to simplify understanding of control flow. Denoting an LED strip as \texttt{L} and the index represented as an integer literal, this example toggles each LED \footnote{in the case of the telos motes, we will consider the LED strip to have an LED\{0, 1, 2\}} successively, separated by a 500ms delay and is repeated indefinitely through a while-loop.

%Within \texttt{Toggle()}, we use the externally realized LED concepts \texttt{Status()} to get the current state of an LED strip, and \texttt{Set()} to set the specified LED at an index to complement. Retrieving the status, as well as setting an LED strip at an index to on or off requires more memory allocation to save these to temporary Boolean and Integer types. In this instance, we note that we need to have the capability of creating and destroying an arbitrary number of times, provides another key motivation for dynamic memory allocation. Determining the minimum memory size needed, becomes increasingly difficult with the while statement. Similarly, other programs may require an implementation of stacks or queues that require a locally unbounded depth.

%Turning to the \texttt{main} function, we first call the appropriate create function to instantiate and allocate all global variables required by the facility. We then make a call to the user supplied \texttt{Main()} operation, which performs the user code. In normal applications the destroy method will be invoked, however here we employ a busy which prevents it from ever being called.

%A wait utility is then called using the \texttt{Std\_Clock\_Fac} facility. This facility uses the externally realized concept to the msp430 \texttt{\_\_delay\_cycles()} command. Internally the MCU is delayed for a given number of cycles.

\begin{figure}
\begin{verbatim}
int Main() {
  while(TRUE) {
    Toggle(L, 0);
    Wait_500_Milli_Seconds();
    Toggle(L, 1);
    Wait_500_Milli_Seconds();
    Toggle(L, 2);
    Wait_500_Milli_Seconds();
  }
}

int main() {
  LED_Telos_Demo_create();
  Main();
  LED_Telos_Demo_destroy();
}
\end{verbatim}
\label{fig:loop}
\end{figure}
%
%
%
%
%
%
%
%
%
%
%
%
%
%

### 2014-05-01 tech-project/paper.tex
changes since initial turn in

\begin{document}
\conferenceinfo{Clemson 2014}{7th Clemson University Mini-Conference on Embedded Network Systems}
\title{An Extensible Approach to Verification of Embedded Network Systems*}

%Re-evaluating Verification of Embedded Network Systems:
%An Extensible Approach to Verification of Embedded Network Systems*}
\numberofauthors{2}
\author{
\alignauthor
%author
Daniel Welch\\
\affaddr{School of Computing, Clemson University}\\
\affaddr{Clemson, South Carolina, 29634}\\
\email{\texttt{\large{\{dtwelch\}@clemson.edu}}}
%author
\alignauthor
Takumi Bolte\\
\affaddr{School of Computing, Clemson University}\\
\affaddr{Clemson, South Carolina, 29634}\\
\email{\texttt{\large{\{tbolte\}@clemson.edu}}}
}
\maketitle

\input{abstract.tex}
\category{D.2.8}{Software Engineering and Data Communication}{Verification}[VCs, automated proving, modular software]
\terms{Reliability, Verification, Languages, Networks}
\keywords{automation, components, formal methods, specification, verifying compiler, embedded networks, wireless sensing}

\input{introduction.tex}
\input{telos_platform.tex}
\input{specify_led.tex}
\input{implementation.tex}
\input{acknowledge.tex}

>>>>

\begin{document}
\conferenceinfo{Clemson 2014}{7th Clemson University Mini-Conference on Embedded Network Systems}
\title{An Extensible Approach to Verification of Embedded Systems}

\numberofauthors{2}
\author{
\alignauthor
%author
Daniel Welch\\
\affaddr{School of Computing, Clemson University}\\
\affaddr{Clemson, South Carolina, 29634}\\
\email{\texttt{\large{dtwelch@clemson.edu}}}
%author
\alignauthor
Takumi Bolte\\
\affaddr{School of Computing, Clemson University}\\
\affaddr{Clemson, South Carolina, 29634}\\
\email{\texttt{\large{tbolte@clemson.edu}}}
}
\maketitle

\input{abstract.tex}
\category{D.2.8}{Software Engineering and Data Communication}{Verification}[VCs, automated proving, modular software]
\terms{Reliability, Verification, Languages, Networks}
\keywords{automation, components, formal methods, specification, verifying compiler, embedded systems, wireless sensing}

\input{introduction.tex}
\input{telos_platform.tex}
\input{specify_led.tex}
\input{implementation.tex}
\input{led_trans.tex}
\input{related_works_and_conclusion.tex}
\input{acknowledge.tex}
